Lab 2: Regression to Study the Spread of Covid-19

### 2020-11-15 lab_2.Rmd
Update lab_2.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r lib,include=FALSE}
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)

@@

```{r variables}
df <- read.csv("covid-19.csv", header = TRUE)

df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
  select(case_rate_100k, population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, mask_public, mask_legal)
head(df)
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r lib,include=FALSE}
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
```

@@

```{r variables}
df <- read.csv("covid-19.csv", header = TRUE)
names(df)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
  select(case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal)
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

@@

model2b <- lm(case_rate_100k ~log(population_density) + I((white_pct)^2) , data = df)
coeftest(model2b, vcov = vcovHC)

For Model 3. Policy related variables will be introduced for the analytics.

df <- df %>% 
  mutate(
    mask_public_bool = case_when(
      mask_public == 0 ~ 0, 
      !(mask_public == 0) ~ 1
      )
  )
head(df)

hist(df$case_rate_100k[df$mask_legal == 1], xlab='Case Rate', ylab='Count', main='Case Rate - Mask Legal')
hist(df$case_rate_100k[df$mask_legal == 0], xlab='Case Rate', ylab='Count', main='Dem Voter - Mask not legal')

hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', main='Case Rate - Mask Mandate in Public')
hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', main='Dem Voter - No Mask Mandate in Public')

model3 <- lm(case_rate_100k ~log(population_density) + I((white_pct)^2) + mask_legal, data = df)
model3 <- lm(case_rate_100k ~ log(population_density) +white_pct+ mask_public_bool, data = df)
#model3 <- lm(death_rate_100k ~ log(population_density) + mask_public_bool, data = df)
coeftest(model3, vcov = vcovHC)

df %>% 
  ggplot(aes(x = mask_public_bool, y = case_rate_100k, color = case_rate_100k)) + 
  geom_point(aes(size=case_rate_100k)) + 
  labs(
    title = 'Case Rate vs Mask Public Mandate', 
    x = 'Mask Mandate in Public', 
    y = 'Case Rate', 
    color = 'Case Rate'
  )

df$close_period = as.integer( df$business_reopen-df$business_closed)
head(df)
hist(df$close_period, xlab='close period', ylab='Count', main='close length')
df %>% 
  ggplot(aes(x = close_period, y = case_rate_100k, color = close_period)) + 
  geom_point() + 
  labs(
    title = 'Case Rate vs Mask Public Mandate', 
    x = 'Mask Mandate in Public', 
    y = 'Case Rate', 
    color = 'Case Rate'
  )

model3 <- lm(case_rate_100k ~ close_period, data = df)
coeftest(model3, vcov = vcovHC)

df[is.na(df$business_closed),]
df[42,'business_closed'] = max(df$business_closed[!is.na(df$business_reopen)])
mean(df$business_closed)
df$business_close_diff = df$business_closed-mean(df$business_closed)
head(df)

df %>% 
  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = close_period)) + 
  geom_point() + 
  labs(
    title = 'Case Rate vs Mask Public Mandate', 
    x = 'Mask Mandate in Public', 
    y = 'Case Rate', 
    color = 'Case Rate'
  )

model3 <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
coeftest(model3, vcov = vcovHC)
